---
title: "Assignment Analysis 1"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
  
date: "2023-02-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Required Packages
```{r}
library(stringr)
library(skimr)
library(tidyverse)

```

### Question 1

```{r cars}

chess = readLines("chess_classic_games.txt")    #Read raw data
#His <- read.csv("Full_Data_Classic_Q1.csv") #Used this to check my solution
                                             #with the correct one
chess_2 = str_split_fixed(chess, " ", 2)     

is_metadata = str_detect(chess, "^\\[.*\\]$") # Separating meta data from moves line

pgn_meta = chess

pgn_meta[!is_metadata] = ""           #Replacing moves line with an empty line

pgn_meta2 = str_split_fixed(pgn_meta, " ", 2)
pgn_meta3= pgn_meta2[!apply(pgn_meta2 == "", 1, all),]  #Removing empty rows

vars = str_replace(pgn_meta3[,1], "\\[", "")   #Erasing [ from metadata
varnames = unique(vars)                        #Finding unique variable names
varnames = varnames[!varnames == '']           #Removing empty variable names
values = pgn_meta3[,2]

values = str_replace_all(values, "\"", "")      #Erasing [ from metadata values

values = str_replace(values, "\\]$", "")        #Erasing ] from metadata values

New_event = str_detect(pgn_meta3[,1], "Event")  #Number of line where there's a
                                                #new Event in cleaned data
New_event_index = which (New_event)
New_event_index = c(New_event_index,nrow(pgn_meta3)) #including last row index
N_games = length(New_event_index)-1    #Number of games played. -1 because of
                                       #the previous code line

MyData <- vector( "character" , 19 )  #Initializing a vector to keep generated rows


for (i in 1:N_games){
  
  if (i==N_games){ #This if condition is because we have to extract last line data differently
    a=New_event_index[i]
    b=New_event_index[i+1]
    
  } else {
    
    a=New_event_index[i]
    b=New_event_index[i+1]-1
    
  }
  
  
  vars2 <- vars[a:b]  #Slicing variable names for the particular row [i]
  reorder_index <- match(varnames,vars2) #matching variable names[i] with unique varnames
  vars2_reordered <-vars2[reorder_index] #Ordering variable names like varnames
  values_reordered <-values[a:b][reorder_index] #ordering values to correspond to varnames
  
  
  values_reordered <- na.omit(values_reordered)  #omitting NA's so they don't mess ordered values
  ind <- varnames %in% vars2_reordered   #Used to know where each value should go
                                       #if values are not ordered as others in metadata
  
  New_data <- vector( "character" , 19 ) #initializing data row
  
  New_data[ind] <- values_reordered 
  MyData <- rbind (MyData, New_data) #Adding new data below the other data
  
}

MyData2= MyData[!apply(MyData == "", 1, all),]   #Erasing the first row because it was blank

My_Df <- data.frame(MyData2)                    #Changing vector to data frame

colnames(My_Df)<-varnames    #Column names should be varnames(unique variable names)
rownames(My_Df)<-c()        



cols.num <- c("WhiteElo","BlackElo","WhiteRatingDiff","BlackRatingDiff")
My_Df[cols.num] <- sapply(My_Df[cols.num],as.numeric)





print(My_Df[6:10,])
 
skim(My_Df)
```
\newpage

### Question 2

```{r}
NEvent = str_detect(chess_2[,1], "Event")     #Number of line where there's a new Event
tag_line = which (NEvent)                     # Finding the index of 

empty_lines <- grep("^$", chess)              # Finding the index of empty lines in raw data
differences <- diff(empty_lines)              
moves_line_ind <- which(differences == 2)#If distance is two, means we have a moves line between them
moves_line2 <- empty_lines[moves_line_ind]+1 # Moves line is one line after the first empty line

#Nmoves = str_detect(chess_2[,1], "1.") 
moves_line = c(moves_line2,nrow(chess_2))     #Adding the last row index for last event's moves line

df2 <- data.frame(moves_line = moves_line,tag_line = tag_line) #Adding columns to the data frame
My_Df2 <- cbind(My_Df,df2)

cols.num <- c("moves_line","tag_line")
My_Df2[cols.num] <- sapply(My_Df2[cols.num],as.numeric)       #Making them numeric values

print(My_Df2[6:10,])


skim(My_Df2)

```

\newpage

### Question 3


There are two approaches for this:
1. Taking the quantiles for Black and White separately: (quantiles_WhiteElo and quantiles_BlackElo)
2. Taking the quantiles for all the players, even recurring players (everybody_quantile)




```{r}
quantiles_WhiteElo <- quantile(My_Df2$WhiteElo,probs=c(0.01,0.05,0.25,0.5,0.75,0.90,0.99,0.999,1))
quantiles_BlackElo <- quantile(My_Df2$BlackElo,probs=c(0.01,0.05,0.25,0.5,0.75,0.90,0.99,0.999,1))
mixed_players <- cbind(My_Df2$BlackElo,My_Df2$WhiteElo)
#The above line has every player's Elo in front of it
everybody_quantile <- quantile(mixed_players, probs=c(0.01,0.05,0.25,0.5,0.75,0.90,0.99,0.999,1))

print(quantiles_BlackElo)
print(quantiles_WhiteElo)
print(everybody_quantile)
```

\newpage


### Question 4

```{r}
#Extracting white player score

Score <- sub("1/2","0.5",sub("\\-.*","",My_Df2$Result) ) #Replacing 1/2 with 0.5

White_Score <- as.numeric(Score) 
Diff_Elos <- My_Df2$WhiteElo - My_Df2$BlackElo #Difference of White and Black Elos

#linear model
White_Score_Df_Elo<-as.data.frame(cbind(White_Score,Diff_Elos))
Model_1 <- lm( White_Score~Diff_Elos , data = White_Score_Df_Elo)

# How many points worse should the first player own to win exactly half of the time?
#First we find the average white score : (1+0)/2 = 0.5 then see what Diff_Elos this correspond to.
Diff_To_win_half <- ((0.5 - Model_1$coefficients[1])/Model_1$coefficients[2])

print(Diff_To_win_half)

summary(Model_1)
```

According to the value "Diff_To_win_half", White should have about 19 less Elos to win exactly half of the time. This proves the slight advantage of playing first (playing with white pieces).

\newpage

### Question 5

```{r}
#Filtering for players with 2000 Elos abd above:
Filtered_Df2 <- filter(My_Df2,WhiteElo >= 2000 & BlackElo >=2000) 

#...And doing the same as the previous question:
Score2 <- sub("1/2","0.5",sub("\\-.*","",Filtered_Df2$Result) )

White_Score2 <- as.numeric(Score2)
Diff_Elos2 <- Filtered_Df2$WhiteElo - Filtered_Df2$BlackElo


White_Score_Df_Elo2<-as.data.frame(cbind(White_Score2,Diff_Elos2))
Model_2 <- lm( White_Score2~Diff_Elos2 , data = White_Score_Df_Elo2)

Diff_To_win_half2 <- ((0.5 - Model_2$coefficients[1])/Model_2$coefficients[2])

print(Diff_To_win_half2)

summary(Model_2)

```

As we can see there's not a significant change in the answer. It is known that professional players playing with black pieces can reduce the advantage of white players with certain moves and strategies. However, the answer "-18.47" is not so different than -19.144.

\newpage

### Question 6

```{r}
first_move_col <- data.frame(c("")) #Intializing column for first move
first_move_ind <- data.frame(c("")) #Intializing column for first move indicator
N <- nrow(My_Df2) #This is number of games

for (j in 1:N){
  clean <- sub(" \\{.*","",chess[My_Df2$moves_line][j]) #using raw data and
  clean <- sub("(.*? .*?) .*", "\\1",clean)
             # moves line indicator to clean first move from the moves line
  
  cleaned = str_split_fixed(clean, " ", 2) #isolating moves line from move number
                                           #(e.g.: 1. from d4)
  moves_df <- data.frame(cleaned)
  if (moves_df$X2 == "e4"){
    
    first_move_ind[j,] = 1  #If first move is e4, indicator shows 1 else, 0.
  } else {
    first_move_ind[j,] = 0
  }
    
  first_move_col[j,] <- moves_df$X2
}

#Adding new column to the data set
df3 <- data.frame(first_move = first_move_col,first_move_ind = first_move_ind)
My_Df3_Q6 <- cbind(My_Df2,df3)

colnames(My_Df3_Q6)[22] <- "first move"
colnames(My_Df3_Q6)[23] <- "first move indicator"

table(My_Df3_Q6$`first move`)
table(My_Df3_Q6$`first move indicator`)
```


\newpage


### Question 7

```{r}

My_Df3_Q7 <- filter(My_Df3_Q6, first_move_ind==1)
freq <- table (findInterval(My_Df3_Q7$WhiteElo,c(-1000000,seq(850,2550,50))))

freq_df <- as.data.frame(freq)
freq_df[,2]
freq_df[,1] <-as.numeric(freq_df[,1])

bin = c("<850","850-900","900-950","950-1000","1000-1050","1050-1100"
        ,"1100-1150","1150-1200", "1200-1250",
        "1250-1300","1300-1350","1350-1400","1400-1450",
        "1450-1500","1500-1550","1550-1600","1600-1650","1650-1700",
        "1700-1750", "1750-1800","1800-1850","1850-1900","1900-1950",
        "1950-2000","2000-2050","2050-2100","2100-2150","2150-2200", "2200-2250",
        "2250-2300","2300-2350","2350-2400","2400-2450","2450-2500","2500-2550")
#freq_df[,1]= bin

plot(freq_df[,1],freq_df[,2],type='b',xlab = "bins",ylab = "Frequencies")
```
 
 Looks like people with an average degree of skill use e4 as their first move more frequently. More professional players might use different strategies and hence might not use e4 as their first move.
On the other hand less skilled players use a variety of moves other than e4 as their first move.
\newpage

### Question 8


```{r}

avgElo <- as.numeric((My_Df2$WhiteElo + My_Df2$BlackElo)/2) #Taking the average
dex = My_Df3_Q6$moves_line  #finding indexes for moves line
count_blunder <- data.frame(c("")) #Initializing column for blunder moves count

for (m in 1:N){
  count_blunder[m,] <- str_count(chess[dex[m]],"\\?") #looking for "?/??/?!/.."
}

Elo_Blunder_count<-as.data.frame(cbind(count_blunder,avgElo))
colnames(Elo_Blunder_count)[1] <- "count_blunder"
Model_3 <- lm(Elo_Blunder_count$count_blunder~avgElo , data = Elo_Blunder_count)

summary(Model_3)

```

\newpage

### Question 9

```{r}
#Extracting time control without the + sign, an adding 1 to avoid ln(0)
T_ctrl <- as.numeric(sub("\\+.*","",My_Df2$TimeControl))+1. 
Model_4 <- lm(Elo_Blunder_count$count_blunder~avgElo+ I(log(T_ctrl)) , data = Elo_Blunder_count)
summary(Model_4)
```

\newpage

### Question 10
```{r}
#Step 1
chess = readLines("chess_960_games.txt")    #Read raw data

chess_2 = str_split_fixed(chess, " ", 2)     

is_metadata = str_detect(chess, "^\\[.*\\]$") # Separating meta data from moves line

pgn_meta = chess

pgn_meta[!is_metadata] = ""           #Replacing moves line with an empty line

pgn_meta2 = str_split_fixed(pgn_meta, " ", 2)
pgn_meta3= pgn_meta2[!apply(pgn_meta2 == "", 1, all),]  #Removing empty rows

vars = str_replace(pgn_meta3[,1], "\\[", "")   #Erasing [ from metadata
varnames = unique(vars)                        #Finding unique variable names
varnames = varnames[!varnames == '']           #Removing empty variable names
values = pgn_meta3[,2]

values = str_replace_all(values, "\"", "")      #Erasing [ from metadata values

values = str_replace(values, "\\]$", "")        #Erasing ] from metadata values

New_event = str_detect(pgn_meta3[,1], "Event")  #Number of line where there's a
                                                #new Event in cleaned data
New_event_index = which (New_event)
New_event_index = c(New_event_index,nrow(pgn_meta3)) #including last row index
N_games = length(New_event_index)-1    #Number of games played. -1 because of
                                       #the previous code line
#Initializing a vector to keep generated rows
MyData <- vector( "character" , length(varnames) )  


for (i in 1:N_games){
  
  if (i==N_games){ #This if condition is because we have to extract last line data differently
    a=New_event_index[i]
    b=New_event_index[i+1]
    
  } else {
    
    a=New_event_index[i]
    b=New_event_index[i+1]-1
    
  }
  
  
  vars2 <- vars[a:b]  #Slicing variable names for the particular row [i]
  reorder_index <- match(varnames,vars2) #matching variable names[i] with unique varnames
  vars2_reordered <-vars2[reorder_index] #Ordering variable names like varnames
  values_reordered <-values[a:b][reorder_index] #ordering values to correspond to varnames
  
  
  values_reordered <- na.omit(values_reordered)  #omitting NA's so they don't mess ordered values
  ind <- varnames %in% vars2_reordered   #Used to know where each value should go
                                       #if values are not ordered as others in metadata
  
  New_data <- vector( "character" , length(varnames) ) #initializing data row
  
  New_data[ind] <- values_reordered 
  MyData <- rbind (MyData, New_data) #Adding new data below the other data
  
}

MyData2= MyData[!apply(MyData == "", 1, all),]   #Erasing the first row because it was blank

My_Df <- data.frame(MyData2)                    #Changing vector to data frame

colnames(My_Df)<-varnames    #Column names should be varnames(unique variable names)
rownames(My_Df)<-c()        



cols.num <- c("WhiteElo","BlackElo","WhiteRatingDiff","BlackRatingDiff","SetUp")
My_Df[cols.num] <- sapply(My_Df[cols.num],as.numeric)

print(My_Df[6:10,])
 

skim(My_Df)

```

\newpage

```{r}

#Step 2
NEvent = str_detect(chess_2[,1], "Event")     #Number of line where there's a new Event
tag_line = which (NEvent)                     # Finding the index of 

empty_lines <- grep("^$", chess)              # Finding the index of empty lines in raw data
differences <- diff(empty_lines)              
moves_line_ind <- which(differences == 2)#If distance is two, means we have a moves line between them
moves_line2 <- empty_lines[moves_line_ind]+1 # Moves line is one line after the first empty line

#Nmoves = str_detect(chess_2[,1], "1.") 
moves_line = c(moves_line2,nrow(chess_2))     #Adding the last row index for last event's moves line

df2 <- data.frame(moves_line = moves_line,tag_line = tag_line) #Adding columns to the data frame
My_Df2 <- cbind(My_Df,df2)

cols.num <- c("moves_line","tag_line")
My_Df2[cols.num] <- sapply(My_Df2[cols.num],as.numeric)       #Making them numeric values

print(My_Df2[6:10,])



skim(My_Df2)
#write.csv(My_Df2, "~/Desktop/UW 2/Data Analysis - Stat 874/A2/Q10.csv", row.names=FALSE)
```

\newpage


### Question 11


Question 6 redone for 960 games:

```{r}

first_move_col <- data.frame(c("")) #Intializing column for first move
first_move_ind <- data.frame(c("")) #Intializing column for first move indicator
N <- nrow(My_Df2) #This is number of games

for (j in 1:N){
  clean <- sub(" \\{.*","",chess[My_Df2$moves_line][j]) #using raw data and
  clean <- sub("(.*? .*?) .*", "\\1",clean)
             # moves line indicator to clean first move from the moves line
  
  cleaned = str_split_fixed(clean, " ", 2) #isolating moves line from move number
                                           #(e.g.: 1. from d4)
  moves_df <- data.frame(cleaned)
  if (moves_df$X2 == "e4"){
    
    first_move_ind[j,] = 1  #If first move is e4, indicator shows 1 else, 0.
  } else {
    first_move_ind[j,] = 0
  }
    
  first_move_col[j,] <- moves_df$X2
}

#Adding new column to the data set
df3 <- data.frame(first_move = first_move_col,first_move_ind = first_move_ind)
My_Df3_Q6 <- cbind(My_Df2,df3) #adding two new column

colnames(My_Df3_Q6)[23] <- "first move"
colnames(My_Df3_Q6)[24] <- "first move indicator" #Renaming columns

table(My_Df3_Q6$`first move`)
table(My_Df3_Q6$`first move indicator`)

```

\newpage

Question 7 redone for 960 games:

```{r}

My_Df3_Q7 <- filter(My_Df3_Q6, first_move_ind==1)
freq <- table (findInterval(My_Df3_Q7$WhiteElo,c(-1000000,seq(850,2550,50))))

freq_df <- as.data.frame(freq)
freq_df[,2]
freq_df[,1] <-as.numeric(freq_df[,1])

bin = c("<850","850-900","900-950","950-1000","1000-1050","1050-1100"
        ,"1100-1150","1150-1200", "1200-1250",
        "1250-1300","1300-1350","1350-1400","1400-1450",
        "1450-1500","1500-1550","1550-1600","1600-1650","1650-1700",
        "1700-1750", "1750-1800","1800-1850","1850-1900","1900-1950",
        "1950-2000","2000-2050","2050-2100","2100-2150","2150-2200", "2200-2250",
        "2250-2300","2300-2350","2350-2400","2400-2450","2450-2500","2500-2550")
bin <- factor(bin)
#freq_df[,1]= bin
#par(las = 2)
plot(freq_df[,1],freq_df[,2],type='b',xlab = "bins",ylab = "Frequencies")

```
