---
title: "Midterm Project"
author: "Tina Hajinejad"
date: "2023-02-24"
output:
  html_document: default
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries:

library(stringr)
library(skimr)
library(tidyverse)
library(lubridate)
library(plyr)
library(dplyr)
library(ggplot2)
library(glmnet)
library(naniar)
library(mice)
library(abind)
library(zoo)
```

## Initializing


```{r Merge Data}

#First Read from csv files
tiketcom = read.csv("tiketcom_bestprice.csv")
distances = read.csv("distance_between_indonesian_airports.csv")

#Now we clean datasets

#Cleaning tiketcom_bestprice.csv
prices_names = str_split_fixed(names(tiketcom), "\\.", 5) 
prices_values = vector("character", 5)

for (i in 1:nrow(tiketcom)){
  val = vector("character", 5)
  val = str_split_fixed(tiketcom[i,1], "\\|", 5)
  prices_values <- rbind(prices_values,val)
}
prices = prices_values[!apply(prices_values == "", 1, all),]   #Erasing the first row because it was blank
prices_df <- data.frame(prices)                    #Changing vector to data frame

colnames(prices_df)<-prices_names    #Column names should be varnames(unique variable names)
rownames(prices_df)<-c()        



```

```{r}

#Cleaning distance_between_indonesian_airports.csv
distance_names = str_split_fixed(names(distances), "\\.", 4) 
distance_values = vector("character", 4)

for (i in 1:nrow(distances)){
  val = vector("character", 4)
  val = str_split_fixed(distances[i,1], "\\|", 4)
  distance_values <- rbind(distance_values,val)
}
distances = distance_values[!apply(distance_values == "", 1, all),]   #Erasing the first row because it was blank
distance_df <- data.frame(distances)                    #Changing vector to data frame

colnames(distance_df)<-distance_names    #Column names should be varnames(unique variable names)
rownames(distance_df)<-c()        
distance_df$distance_km <- as.numeric(distance_df$distance_km)

```




```{r pressure, echo=FALSE}

distance_km <- vector("numeric", length = nrow(prices_df))
flight_time_hour <- vector("numeric", length = nrow(prices_df))
for (i in 1:nrow(prices_df)){
  for (j in 1:nrow(distance_df)){
    ind = str_detect(prices_df$destination[i] , distance_df$airport_to[j])
    if (ind == TRUE){
      distance_km[i] <- distance_df$distance_km[j]
      flight_time_hour[i] <- distance_df$flight_time_hour[j]
      break
    }
  }
}

prices_df$distance_km <- distance_km
prices_df$flight_time_hour <- flight_time_hour


```


```{r}
#Changing types:
cols.num <- c("best_price" , "distance_km" , "flight_time_hour")
prices_df[cols.num] <- sapply(prices_df[cols.num],as.numeric) 
skim(prices_df)
```


## Question 2

```{r Plot}

# Adding a column to see what day is each date

days <- wday(prices_df$depart_date, label = TRUE)
days_num <- wday(prices_df$depart_date)

#Making a new df

price_day <- cbind(prices_df, days, days_num)

#Find the average of price ticket based on 

price_stats = ddply(price_day, "days_num", summarize,
                     avg_pl = mean(best_price))



plot(price_stats[,1],price_stats[,2],type ="l",xlab = "days",ylab = "avg price")
```

```{r}
#Test: See what flies happen on mondays.
#Flights each day of the week
mon = filter(price_day, days == "Mon")
mon_type = unique(mon$destination)


tue = filter(price_day, days == "Tue")
tue_type = unique(tue$destination)
####
####
# Flight pattern for a destination / each day of week
btj <- filter(price_day, destination == "BTJ")
btj_stats = ddply(btj, "days_num", summarize,
                     avg_pl = mean(best_price))

plot(btj_stats[,1],btj_stats[,2],type ="l",xlab = "days",ylab = "avg price")




```

\newpage

We will now see how average price of plane tickets change based on the departure date.
```{r}

### Prices as a function of departure date 
price_day$depart_date<-as.Date(price_day$depart_date)
price_stats_date = ddply(price_day, "depart_date", summarize,
                     avg_p_d = mean(best_price))
plot(price_stats_date[,1],price_stats_date[,2],type ="l",xlab = "days",ylab = "avg price")


```

Nothing is visible from this graph excpt that price tickets for summer are high because everybody is going on a vacation but prices for 2024 are now low because they are far off in the future.

Now let's see if flights destinations have a say in this:

```{r}
price_day$depart_date<-as.Date(price_day$depart_date)
price_stats_date_2 = ddply(price_day, .(depart_date,destination), summarize,
                     avg_2 = mean(best_price))
#plot(price_stats_date[,1],price_stats_date[,2],type ="l",xlab = "days",ylab = "avg price")


ggplot(price_stats_date_2, aes(x = depart_date, y = avg_2, color = destination)) +
  geom_line() +
  labs(x = "Departure Date", y = "Average Price", color = "Destination") +
  theme_bw() +
  theme(aspect.ratio=1/2) #Long and skinny


```

This isn't helping. I try to separate types of flights, based on flight time
```{r}

# Finding the point where flight_hours more than that is considered "long flight"
# and flights wih durations less than that is considered "short" flight
distance_df$flight_time_hour <- as.numeric(distance_df$flight_time_hour)
separator <- mean(distance_df$flight_time_hour)

short_flight = filter(price_day, flight_time_hour < separator )
long_flight = filter(price_day, flight_time_hour > separator )

# Finding averages for short and long flights, then plotting

short_flight_stats = ddply(short_flight, "depart_date", summarize,
                     avg_sf = mean(best_price))
long_flight_stats = ddply(long_flight, "depart_date", summarize,
                     avg_lf = mean(best_price))


# Combine data into a data frame
#combine_df <- cbind(short_flight_stats, long_flight_stats$avg_lf)

# Plot both curves
ggplot() +
  geom_line(data = short_flight_stats, aes(x = depart_date, y = avg_sf, color = "Short Flights")) +
  geom_line(data = long_flight_stats, aes(x = depart_date, y = avg_lf, color = "Long Flights")) +
  labs(x = "Departure Date", y = "Average Price", title = "Average Price for Short and Long Flights") +
  scale_color_manual(values = c("Short Flights" = "blue", "Long Flights" = "red")) +
  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "100 day")


```


```{r Q3}
dist_seq = seq(from=130, to =4000, by=100)

my_bin <- cut(prices_df$distance_km, dist_seq)

my_dist <- data.frame(prices_df, my_bin)

dist_price_bin = ddply(my_dist, "my_bin", summarize,
                     avg_bin = mean(best_price))

dist_price_bin_all <- complete(dist_price_bin, my_bin)
dist_price_bin_all$my_bin <- as.factor(dist_price_bin_all$my_bin)
ggplot(dist_price_bin_all, aes(x = my_bin, y = avg_bin)) +
  geom_point(size = 2) +
  xlab("Distance bin") +
  ylab("Average price") +
  ggtitle("Average price by distance bin") +
  theme(axis.text.x = element_text(angle = 45, size = 8)) +
  scale_x_discrete(breaks = dist_price_bin_all$my_bin[c(T, F, F)])
```

```{r}
#Q4a 5pts) Plot the number of flights by calendar day (of departure)

my_data <- price_day %>% mutate(ones = 1)

n_flights_df = ddply(my_data, "depart_date", summarize,
                     n_flights = sum(ones))


#dn_flights_df_all <- complete(dist_price_bin, my_bin)

ggplot(n_flights_df, aes(x = depart_date, y = n_flights)) +
  geom_point(size = 1) +
  xlab("Depart Date") +
  ylab("Number of flights") +
  ggtitle("Number of flights for each date") +
  theme(axis.text.x = element_text(angle = 45, size = 10))

```

There are so many missing data. Let's complete them with NAs:
```{r}
depart_date <- seq(from = as.Date("2023-01-01"), to = as.Date("2024-06-26"), by = "day")
depart_date <- data.frame(depart_date)

new_data <- n_flights_df %>% 
  right_join(depart_date, by = "depart_date") %>% 
  mutate(n_flights = ifelse(is.na(n_flights), NA, n_flights))

sorted_data <- new_data %>%
  arrange(depart_date)

vis_miss(sorted_data)


```
We can see many many flights information are missing(mainly between 2023-08 to 2023-12)
and many others in 2024.
Now that we have added NAs to the data, let's impute them.:
```{r}
#Q4b) imputation
#we have this sorted dataframe that shows the number of flights per departure date. 
#I added the dates with no flights and filled them with NAs, because that's necessary for imputation.
#There are several ways for imputation.
#Let's try: Remove cases with missing values for only the variables used. Because there are so many of them I don't prefer this:
#We can check this by:

vis_miss(sorted_data)


```
Now let's Check : Multiple imputation.
```{r}
# Make 5 imputations:
tempData <- mice(sorted_data,m=5,maxit=10,method='pmm',seed=500)
summary(tempData)
completedData <- complete(tempData,1)
summary(completedData)
xyplot(tempData,n_flights ~ depart_date, pch=18, cex=0.8)

```
Another option is The zoo package (Z’s Ordered Observations) is mostly for time series data, where information is time-stamped or otherwise has a definite time-like ordering to it.
na.approx(), which interpolates NA values in a set of ordered values.

```{r}
z = zoo(sorted_data[,2],sorted_data[,1])
z_complete <- na.approx(z)

#now let's plt:
```



```{r}

#Q5

cols.num <- c("days_num","flight_time_hour","best_price","distance_km")
price_day[cols.num] <- sapply(price_day[cols.num],as.numeric)

#I use price_day data frame.

#First, we need to the define response variable

price_day$destination <- as.factor(price_day$destination)
price_day

y <- price_day$best_price
# and also define a matrix of predictor variables
x <- data.matrix(price_day[, c('days_num','flight_time_hour')])

# Find best lambda
cv_model <- cv.glmnet(x, y, alpha = 1)

best_lambda <- cv_model$lambda.min
best_lambda
```

```{r}
summary(price_day[, c('distance_km', 'flight_time_hour','days_num')])

# standardize the predictors
x <- as.matrix(scale(price_day[, c('distance_km', 'flight_time_hour','days_num')]))
y <- price_day$best_price

# fit the LASSO model
lasso_model <- cv.glmnet(x, y, alpha = 1)
best_lambda <- lasso_model$lambda.min
best_lambda
best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```
